### 2018-05-02 vignette.Rmd
temp pushing bits and pieces

---
title: "An introduction to pglmm in lme4"
author: "Michael Li and Ben Bolker"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
output: pdf_document
---

# 1 Preamble

You can install lme4 by the developer version on github using devtools ``devtools:::install_github(...)''

# Phylogenetic Generalised Linear Mixed Models

What follows is a simple example of how to simulate, and then analyse, the most basic kind of PGLMM.

## Simulate data

```{r}
### simulate phylogenetic tree

library(ape)
## library(MASS)  ## for mvrnorm() ## don't load so we don't screw up dplyr::select()
seed <- 101
set.seed(seed)

# Basic parameters
nspp <- 3

# # fixed effects
beta0 <- 0
beta1 <- 0

# random effects and residual variance (set to zero for binary data)
sd.B0 <- 1
sd.B1 <- 2
sd.resid <- 10
rho.B01 <- 0

# Simulate phylogeny
phy <- rtree(n = nspp)
phy <- compute.brlen(phy, method = "Grafen", power = 0.5)

Vphy <- vcv(phy)

iD <- t(chol(Vphy))
Xsd <- 1

# Generate environmental variable 
x <- rnorm(n=nspp,sd=Xsd)

cormat <- matrix(c(1,rho.B01,rho.B01,1),2,2)
sdvec <- c(sd.B0,sd.B1)
varmat <- sdvec %*% t(sdvec)
covmat <- varmat * cormat

#print(covmat)

Sigma <- kronecker(covmat,Vphy)

#if((signal.B0==signal.B1) & (signal.B0 == FALSE)){
#	Sigma <- kronecker(covmat,diag(nspp))
#}

b.all <- MASS::mvrnorm(n=1
	, mu=rep(c(beta0,beta1),each=nspp)
	, Sigma=Sigma
)

b0 <- b.all[1:nspp] 
b1 <- b.all[(nspp+1):(2*nspp)]

y <- matrix(outer(b0, array(1, dim = c(1, 1))), nrow = nspp,
	ncol = 1) + matrix(b1*x, nrow = nspp, ncol = 1)

e <- rnorm(nspp, sd = sd.resid) # add residual variance

Y <- y + matrix(e, nrow = nspp, ncol = 1)

# name the simulated species 1:nspp and sites 1:nsites
rownames(Y) <- 1:nspp

X <- matrix(x,nrow=nspp,ncol=1)

sp <- matrix(kronecker(matrix(1, nrow = 1, ncol = 1), 1:nspp),
             nrow = nspp, ncol = 1)

dat <- data.frame(Y = Y, X=X, sp = as.factor(sp))

print(dim(dat))

```

## Species branch matrix

- call this SBmat

```{r}

library(Matrix)
library(lme4)
library(dplyr)

source("phyloglmm_setup.R")

phyZ <- phylo.to.Z(phy)

print(phyZ)

```

## Fitting phyloglmm

- I kind of want to keep the orginal name (lmer and glmer).

```{r}

source("new_phylo_setup.R")

dat <- (dat %>% mutate(obs = sp))

## everything can be reduce down to formula,data, and a phy object

fitmod <- phylo_lmm2(Y ~ X + (1 |sp) + (1 + X|sp)
			, data=dat 
			, phylonm = "sp" # wishy-washy
			, phyloZ = phyZ # Maybe we just want the phy object and compute this internally
			, nsp = nspp # make default check
			, control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore")  ## make default
			, REML = FALSE # make default 
)

summary(fitmod)

```

## Simulating using lme4's simulation function

- first fit the object and hack the parameters