### 2019-04-10 example.rmd
done

---
title: "Example"
author: "Michael Li and Ben Bolker"
output:
  pdf_document: default
  mathjax: default
  html_document: null
---

We will reproduce the examples in [chapter 11](http://www.mpcm-evolution.com/practice/online-practical-material-chapter-11/chapter-11-1-simple-model-mcmcglmm) of Garamszegi (ed.) using lme4 and glmmTMB.

To fit we need a random effect in the formula that includes a `(...|phylo)` term, to build the basic random effect structure multipled by the phyloZ matrix (see vignette for more details on phyloZ matrix).

```{r pkgs,message=FALSE, warning=FALSE}
library(ape)
library(Matrix)
library(lme4)
library(MCMCglmm)
library(MASS) 
library(pez)
library(glmmTMB)
library(dplyr)
library(coda)
library(lattice)
library(broom)
library(dotwhisker)
```

```{r phylocode}
source("lme4_phylo_setup.R")
source("glmmTMB_phylo_setup.R")
```

### Get data

From [chapter 11](http://www.mpcm-evolution.com/practice/online-practical-material-chapter-11/chapter-11-1-simple-model-mcmcglmm) of Garamszegi (ed.): data are [here](http://mpcm-evolution.com/OPM/Chapter11_OPM/data.zip)

```{r getdat}
if (!file.exists("data/phylo.nex")) {
    dir.create("data")
    download.file("http://mpcm-evolution.com/OPM/Chapter11_OPM/data.zip",
                  dest="data/OPM_ch11_data.zip")
    setwd("data")
    untar("OPM_ch11_data.zip")
    setwd("..")
}
phylo <- read.nexus("data/phylo.nex")
```

Compute appropriate $Z$ matrix up front, to measure speed (also
reusable in a few places below):

```{r calc_phyloZ,cache=TRUE}
system.time(phyloZ <- phylo.to.Z(phylo))
```

## Result comparison with Gaussian example in chapter 11

```{r G_results}
datG <- read.table("data/data_simple.txt",header=TRUE)
datG$obs <- factor(seq(nrow(datG)))
datG <- datG %>% mutate(sp = phylo)
phylo_lmm_fit <- phylo_lmm(phen~cofactor+(1|sp)
  , data=datG
  , phylonm = "sp"
  , phylo = phylo
  , phyloZ=phyloZ
  , REML = TRUE
  , control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore")
)

print(summary(phylo_lmm_fit))

```

## Result comparison with Gaussian with repeated measures example in chapter 11

```{r GR_results}
datR <- read.table("data/data_repeat.txt",header=TRUE)
datR$obs <- factor(seq(nrow(datR)))
datR <- (datR 
  %>% mutate(sp = species
        , animals = phylo
        )
)
datR$spec_mean_cf <- sapply(split(datR$cofactor,datR$phylo),mean)[datR$phylo]
datR$within_spec_cf <- datR$cofactor-datR$spec_mean_cf
phylo_lmm_fit <- phylo_lmm(phen~spec_mean_cf+within_spec_cf+(1|sp) + (1|animals)
  , data=datR
  , phylonm = "sp"
  , phylo = phylo
  , phyloZ=phyloZ
  , REML = FALSE
  , control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore")
)

print(summary(phylo_lmm_fit))

```

## Result comparison with non-Gaussian example in chapter 11

```{r P_results}
dat <- read.table("data/data_pois.txt",header=TRUE)
dat$obs <- factor(seq(nrow(dat)))

dat <- dat %>% mutate(sp=phylo) 
phylo_glmm_fit <- phylo_glmm(phen_pois~cofactor+(1|sp)+(1|obs)
  , data=dat
  , phylonm = "sp"
  , family = poisson
  , phylo = phylo
  , phyloZ=phyloZ
  , control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore")
)

summary(phylo_glmm_fit)

```