### 2022-08-09 table1.tex

\begin{table}[]
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Model} & \textbf{Method} & \textbf{Data} & \textbf{Platform} \\
\hline
\multirow{2}{*}{
  \begin{tabular}[c]{@{}l@{}}
    Generalized Linear \\ Model (GLM)
\end{tabular}} &
Correlated residual &
\begin{tabular}[c]{@{}l@{}} Single observation \\ per species \end{tabular} &
\begin{tabular}[c]{@{}l@{}}\pkg{nlme}:gls, \\ \pkg{ape}:pic \end{tabular} \\
\cline{2-4} &
\begin{tabular}[c]{@{}l@{}}Residual \\ + phylogenetic intercept\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Single observation \\ per species \end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Pagel's $\lambda$\\ Blomberg's $K$ \\ via \pkg{nlme}:gls\\ \pkg{phylolm}\end{tabular} \\
  \hline 
  \multirow{2}{*}{
    \begin{tabular}[c]{@{}l@{}}Generalized Linear \\ Mixed Model (GLMM)\end{tabular}} &

  \multirow{2}{*}{Random effect}
  & \begin{tabular}[c]{@{}l@{}}Single observation \\per species, \\ Balanced design\end{tabular} &
      \pkg{pez}                                                     \\ \cline{3-4} 
      &  & Unrestricted   &
      \begin{tabular}[c]{@{}l@{}}  \pkg{\plmefour}, \\ \pkg{\pgTMB}, \\ \pkg{phyr} \end{tabular}
      \\
      \hline
         \multirow{2}{*}{Bayesian GLMM}
         & \multirow{2}{*}{Random effect}
         & Balanced design                                                              & \pkg{MCMCglmm}                                                                                       \\ \cline{3-4} 
                                                                                                   &                                                                       & Unrestricted                                                                 & \pkg{brms}                                                                                           \\ \hline
\end{tabular}
\end{table}